Hidden Talents in Context: Multiverse analysis repository

### 2022-01-28 manuscript/manuscript.Rmd
CD revision 1, integrate coauthor feedback

Youth living in adverse conditions tend to score lower on cognitive tests than youth in supportive conditions. For example, youth raised in poverty are more likely to have difficulties learning in school and score lower on standardized tests [@farah2006; @hackman2010; @noble2007; @ursache2016a]. Harsh and unpredictable environments are thought to cause high levels of chronic stress, which can impair brain regions responsible for memory and learning [@blair2012]. Thus, the contemporary view is that exposure to adversity undermines cognitive development and leads to a host of cognitive deficits [see @duncan2017; @hackman2010; @mclaughlin2019; @pechtel2011 for reviews]. Although it is clear that harsh and unpredictable environments are harmful to youth, there is growing interest in the *hidden talents approach*, which proposes that some abilities may be enhanced by adversity, especially skills and abilities that help youth meet real-world challenges in adverse environments [@ellis2017; @ellis2020; @frankenhuis2020; @frankenhuis2013].

The hidden talents framework assumes that development shapes the skills and abilities of individuals in relation to their environment [see @ellis2017; @ellis2020; @frankenhuis2020; @frankenhuis2013 for reviews] and assumes development tailors skills and abilities in relation to their environment. Accordingly, exposure to adversity should shape skills and abilities well-suited for dealing with the challenges associated with harsh and unpredictable environments.However, because specific dimensions of adversity (violence versus resource-scarcity) may pose unique challenges [@ellis2020; @frankenhuis2020; @mclaughlin2019; @sheridan2014], the hidden talents approach proposes that cognitive skills should match specific dimensions of adversity [@frankenhuis2020]. For instance, some evidence suggests physically abused youth can detect angry facial expressions quickly and accurately [@gibb2009; @pollak2008; @pollak2009], people living in poverty have more accurate social perception [@bjornsdottir2017; @dietze2021; @piff2018], and anxiously attached people can more accurately detect deception [@eindor2014; @shoda2013] to list a few examples [see @ellis2017; @ellis2020 for reviews].

Although some abilities themselves are more ecologically relevant to specific dimensions of adversity, hidden talents may also emerge under ecologically relevant conditions. For example, previous studies of hidden talents have tested cognition under ecologically relevant *contexts and situations*. In particular, studies have experimentally manipulated the current context to be uncertain and mildly stressful (or not) and measured performance on executive functioning and working memory tasks [@mittal2015; @young2018]. Strikingly, people from unpredictable environments showed more efficient switching on attention shifting tasks and more accurate responses on working memory updating tasks, but this effect only emerged under conditions of uncertainty. There was no performance increase under control conditions. However, some studies have found similar patterns without uncertainty [see @fields2021 for attention shifting and @nweze2021 for working memory].

Hidden talents may also emerge when testing *content and stimuli* are more ecologically relevant and concrete. Typical cognitive tests use abstract content that is largely detached from the real world. Such abstract materials may disadvantage youth with limited exposure to formal education. Cultural psychologists and anthropologists have demonstrated this fact across the world. For example, people from non-Western cultures tend to score lower than people in Western cultures on standard versions of cognitive tests, but this performance gap closes or disappears with more ecologically relevant versions of the same tests [@banerjee2017; @greenfield2014; @ogbu1981; @richardson1991; @schliemann2002]. This work challenges the assumption that "baseline" performance is best captured by standard/abstract tests. In fact, they suggest that abstract stimuli, and more broadly standard testing conditions, are specific contexts that may not match one's lived experience [@doebel2020]. These ideas also pose an important question: do adversity-exposed youth score higher on tests with ecological stimuli relative to those with abstract stimuli?

One way to test this question is to replace sterile testing materials with ecologically relevant content from the real world. For example, one study tested memory and reasoning among young adults who varied in their exposures to violence [@frankenhuis2020a]. They created two versions of a memory and reasoning task, one neutral (chronological age) and one more ecologically relevant (social dominance). They found that people with more *current* exposures to violence were equally, or more, accurate when memorizing social-dominance relationships than people with fewer exposures. Notably, such violence exposures were associated with less accurate memory for neutral (age) relationships. This pattern was specific *current* exposures: people with more *childhood* exposures to violence memorized both content types less accurately. There were no associations between violence exposures and reasoning performance. Overall, this study suggests that ecologically relevant testing materials may in some conditions 'level the playing field' for adversity-exposed youth [@ellis2020; @frankenhuis2020; @rifkin-graboi2021; @vantassel-baska2018].

Together, research testing performance in ecologically relevant contexts/situations and using ecologically relevant materials have generated innovative ideas and findings. Nonetheless, it remains unclear whether or how these ideas and findings can be practically applied. For example, narrow abilities, such as memorizing social dominance hierarchies or accurately detecting anger in faces, seem to have limited practical value in educational contexts. In addition, although executive functions and working memory are related to many educational outcomes, it is unclear how schools and teachers might leverage skills and abilities that specifically depend on inducing uncertainty. That seems impractical, at best, and unethical, at worst.

Here we address this issue by evaluating the role of ecologically relevant content in attention-shifting and working memory tasks. We define ecological relevant content as testing material that resemble the local environment and therefore have more meaning and consequence. We used two criteria for stimulus selection. First, we selected stimuli that youth would encounter and interact with on a daily basis. Second, we selected stimuli believed to be more consequential or salient to adversity-exposed youth. For example, money is more consequential and perhaps salient to adversity-exposed youth than other neutral objects (a book, dishes, etc.). In addition, we target both between-subjects and within-subject improvement as patterns of interest. That is, improved performance compared to oneself (within-subjects -- one's performance on abstract compared to ecologically task versions), compared to others (between-subjects -- high versus low adversity), or some combination, are each theoretically important patterns of improvement. Our specific expectation was that all youth would score higher on tasks with ecological stimuli. However, we expected this increase to be larger among youth living in adverse conditions compared to youth in supportive conditions.

@@

We then tested performance on two widely used tasks: attention shifting and working memory updating. There were two versions of each task. The first used standard abstract stimuli and the second replaced these stimuli with more ecological stimuli from the real-world (see Figure 1). We then tested the interactive effect of task content and adversity exposure. As such, our primary analyses involved two preregistered confirmatory analyses and one exploratory analysis. Our confirmatory analyses focused on environmental unpredictability and violence exposure. Both dimensions have been associated with improved performance under ecologically relevant conditions (contexts and testing materials) and represent specific environmental challenges. Our exploratory analyses focused on exposure to poverty. Poverty is both relevant to hidden talents (e.g., hidden talents could develop in response to specific challenges of poverty) and a general proxy to adversity rather than a specific exposure.

@@

The current study collected data from `r txt1_sam$raw$n$total` youth (*M~age~* = `r txt1_sam$raw$age$mean`, *SD~age~* = `r txt1_sam$raw$age$sd`) from two sources across Salt Lake City, Utah. The first was a middle school and the second were 5 after-school clubs. The middle school sample (seventh and eighth graders, *N* = `r txt1_sam$raw$n$wjms`; `r txt1_sam$raw$sex$wjms$females` females) formed `r (txt1_sam$raw$n$wjms/txt1_sam$raw$n$total * 100) %>% round(2)`% of the total sample and is socioeconomically and ethnically diverse: `r data1_raw %>% filter(race == 1) %>% nrow()` youth identified as Asian, `r data1_raw %>% filter(race == 2) %>%nrow()` as Native American, `r data1_raw %>% filter(race == 3) %>% nrow()` as African American, `r data1_raw %>% filter(race == 4) %>% nrow()` as White, `r data1_raw %>% filter(race == 5) %>% nrow()` as Hispanic, `r data1_raw %>% filter(race == 6)  %>% nrow()` as Pacific Islander, `r data1_raw %>% filter(race == 7) %>% nrow()` with multiple racial identities, and `r data1_raw %>% filter(is.na(race)) %>% nrow()` were missing. According to school district records, `r (txt1_sam$raw$free_lunch$wjms/txt1_sam$raw$n$wjms * 100) %>% round(2)`% of students were receiving economic assistance based on family income and/or homelessness (i.e., free or reduced-price lunch or fee waivers). The clubs (*N* = `r txt1_sam$raw$n$bgc`; `r txt1_sam$raw$sex$bgc$female` females) contained a wider age range (*M~age~*= `r txt1_sam$raw$age$bgc$mean`, *SD~age~* = `r txt1_sam$raw$age$bgc$sd`). `r (txt1_sam$raw$free_lunch$bgc/txt1_sam$raw$n$bgc * 100) %>% round(2)`% were receiving free or reduced-price lunch and were also ethnically diverse.

@@

Our measure of environmental unpredictability integrated information from the interview and questionnaire. The interview contained three variables of interest. The first measured the level of exposure to living in a biologically non-intact family from birth to age 13. Interviewers asked participants about their family composition and the adults that lived with them in the home. If a participant's family composition changed, for example, because of a divorce, interviewers recorded the ages when participants experienced the change and for how long it persisted. A family was considered non-intact if the participant did not live with both biological parents during any period from birth to age 13. Based on the earliest age of disruption, we indexed the onset of exposure to a non-intact family by creating a four-level variable from 0 to 3. A 0 indicated the participant lived in an intact family (lived with both biological parents) from birth until age 13 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 0) %>% sum(na.rm=T)`); a 1 indicated that a disruption occurred after the age of 5 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 1) %>% sum(na.rm=T)`); a 2 indicated that a disruption occurred before age 5 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 2) %>% sum(na.rm=T)`); and a 3 indicated the participant was born into a non-intact family (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 3) %>% sum(na.rm=T)`). `r is.na(data2_non_arbitrary$unp_interview_disrupted) %>% sum(na.rm=T)`. `r is.na(data2_non_arbitrary$unp_interview_disrupted) %>% sum(na.rm=T) %>% english::as.english() %>% str_to_title()` participants had missing information.

@@

We assessed each participant's poverty exposure using three information sources: interview-assessed parent education and occupational prestige, school or club-provided economic information, and perceived resource scarcity.

@@

```{r table2, tab.id = "table2", tab.cap.style = "Table Caption", tab.cap = "Median standardized effects, 95% confidence intervals, percent significant, and bootstrapped p-values for the main effect of adversity and adversity by task version interactions."}
table.02.0 %>% 
  ungroup() %>% 
  mutate(iv = factor(iv, c("Unpredictability","Violence","Poverty"))) %>%
  arrange(dv,iv) %>% 
  select(-dv) %>% 
  add_column(.after = 5,"empty1" = "") %>% 
  add_row(.before = 1, "iv" = "Attention Shifting") %>% 
  add_row(.after = 4, "iv" = "Working Memory Updating") %>% 
  flextable() %>% 
  merge_h_range(i = 1, j1 = 1, j2 = 10) %>% 
  merge_h_range(i = 5, j1 = 1, j2 = 10) %>% 
  bold(i = c(1,5), j = 1) %>% 
  padding(i = c(2:4, 6:8), j = 1, padding.left = 10) %>% 
  width("empty1", width = .2) %>% 
  set_header_labels(iv = "", empty1 = "", ci_main = "95% CI", ci_int = "95% CI") %>% 
  add_header_row(
    values = c(" ", "Main Effect", " ", "Interaction"),
    colwidths = c(1, 4, 1, 4)
  ) %>% 
  compose(i = 2, j = c(2,7), as_paragraph("\U1D6FD"), part = "header") %>% 
  compose(i = 2, j = c(4,9), as_paragraph(as_i("p "), "(%)"), part = "header") %>% 
  compose(i = 2, j = c(5,10), as_paragraph(as_i("p")), part = "header") %>% 
  align(i = 1:2, align = "center", part = "header") %>% 
  border_remove() %>% 
  border(i = 1, border.top = fp_border_default(), part = "header") %>% 
  border(i = 1, j = c(2:5,7:10), border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 2, border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 8, border.bottom = fp_border_default(), part = "body") %>% 
  bold(i = 1:2, part = "header") %>% 
  set_table_properties(width = 1, layout = "autofit") %>% 
  add_footer_row(
    values = " ",
    colwidths = 10
  ) %>% 
    add_footer_row(
    values = " ",
    colwidths = 10
  ) %>% 
  compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "The p (%) column reflects the number of analyses that produced p-values < .05 out of all analyses for a given multiverse analyses. The total number of analyses for each measure was 64. Overall p-values were computed using a bootstrapping resampling method and reflect the probability of obtaining an effect size as extreme or more extreme given the median effect is 0."), 
    part = "footer"
  )
```

@@

```{r table3, tab.id = "table3", tab.cap.style = "Table Caption", tab.cap = "Median simple effects (unstandardized) and percent significant for task version and adversity."}
table.03.0 %>% 
  ungroup() %>%
  mutate(iv = factor(iv, c("Unpredictability","Violence","Poverty"))) %>%
  arrange(dv,iv) %>% 
  select(-dv) %>% 
  add_column(.after = 5,"empty1" = "") %>% 
  add_row(.before = 1, "iv" = "Attention Shifting") %>% 
  add_row(.after = 4, "iv" = "Working Memory Updating") %>% 
  flextable() %>% 
  merge_h_range(i = 1, j1 = 1, j2 = 10) %>% 
  merge_h_range(i = 5, j1 = 1, j2 = 10) %>% 
  bold(i = c(1,5), j = 1) %>% 
  padding(i = c(2:4, 6:8), j = 1, padding.left = 10) %>% 
  width("empty1", width = .2) %>% 
  set_header_labels(iv = "", empty1 = "") %>% 
  add_header_row(
    values = c(" ", "Task Version", " ", "Adversity Exposure"),
    colwidths = c(1, 4, 1, 4)
  ) %>% 
  compose(i = 2, j = c(2), as_paragraph(as_i("b "), "(abstract)"), part = "header") %>% 
  compose(i = 2, j = c(4), as_paragraph(as_i("b "), "(ecological)"), part = "header") %>% 
  compose(i = 2, j = c(7), as_paragraph(as_i("b "), "(low)"), part = "header") %>% 
  compose(i = 2, j = c(9), as_paragraph(as_i("b "), "(high)"), part = "header") %>% 
  compose(i = 2, j = c(3,8), as_paragraph(as_i("p "), "(%)"), part = "header") %>% 
  compose(i = 2, j = c(5,10), as_paragraph(as_i("p "), "(%)"), part = "header") %>% 
  align(i = 1:2, align = "center", part = "header") %>% 
  border_remove() %>% 
  border(i = 1, border.top = fp_border_default(), part = "header") %>% 
  border(i = 1, j = c(2:5,7:10), border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 2, border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 8, border.bottom = fp_border_default(), part = "body") %>% 
  bold(i = 1:2, part = "header") %>% 
  fontsize(size = 11, part = "body") %>%
  set_table_properties(width = 1, layout = "autofit") %>% 
  add_footer_row(
    values = " ",
    colwidths = 10
  ) %>% 
    add_footer_row(
    values = " ",
    colwidths = 10
  ) %>% 
  compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "Simple effects for Task Version reflect the median effect of adversity in for the abstract version (abstract) and ecological version (ecological). Simple effects for Adversity reflect the effect of task version when adversity exposure is low (-1 SD) and high (+1 SD). P-values < .05 out of all analyses for a given multiverse analyses. The total number of analyses for each measure was 64."), 
    part = "footer"
  )
```

@@

For this expanded multiverse approach, we retained all original data processing decisions (i.e., original 64 combinations of decisions) and added adversity component as an additional decision. For unpredictability, we analyzed each unpredictability component and the composite measure (composite analyses are redundant with primary analyses but were retained for comparison purposes), which resulted in 384 analyses (64 original × three versions of unpredictability × two performance measures (attention shifting and working memory). We applied the same approach to violence exposure (three violence variables, 384 analyses) and poverty exposure (four poverty variables, 512 analyses). The full results of these analyses are reported in the supplement (see SI Table 2).

@@

Leveraging multiverse analysis [@simonsohn2020; @steegen2016], we tested interactions between dimensions of adversity and task version on attention shifting and working memory updating performance. We did not find interaction effects for attention shifting. However, youth exposed to unpredictability and violence showed lower switch costs on average, suggesting slightly improved attention shifting performance irrespective of task content. These findings are consistent with a recent study showing that higher caregiving instability was associated with improved attention shifting [@fields2021]. However, our findings also partially diverge from prior work [@mittal2015; @nweze2021]. For example, whereas we found unpredictability was associated with improved attention shifting on average, Mittal and colleagues [-@mittal2015] found no difference in performance under neutral conditions. And, although they found improved attention shifting for those exposed to high unpredictability under uncertainty, it remains unclear how this effect relates to those obtained in the current study. In addition, Nweze and colleagues [-@nweze2021] found no association between adversity and attention shifting, but they tested institutionalized children rather than measuring adversity dimensions. In general, we should interpret cautiously. Although we followed field standards for calculating switch costs by computing difference scores, such scores ignore participant-level reaction times. Mixed-modeling approaches can account for both overall reaction times and differences across switch and repeat trials, but require more statistical power to test three-way interactions. Future research should consider modeling approaches that capture participant-level reaction times and differences across conditions.

@@

Overall, inconsistencies across prior work and the current attention-shifting and working memory updating findings could be attributed to differences in sampling (adults vs. youth), the sampled range of adversity exposure (restricted vs. broad), or methods/design (within- vs. between-subjects design; lab vs. community/school settings). Future research should both address these inconsistencies and conduct more direct replications of individuals studies to converge on reliable patterns.

>>>>

Youth living in adverse conditions tend to score lower on cognitive tests than youth in supportive conditions. For example, youth raised in poverty are more likely to have difficulties learning in school and score lower on standardized tests [@farah2006; @hackman2010; @noble2007; @ursache2016a]. Harsh and unpredictable environments are thought to cause high levels of chronic stress, which can impair brain regions responsible for memory and learning [@blair2012]. Thus, the contemporary view is that exposure to adversity undermines cognitive development and leads to a host of cognitive deficits [reviewed in @duncan2017; @hackman2010; @mclaughlin2019; @pechtel2011]. Although it is clear that harsh and unpredictable environments are harmful to youth, there is growing interest in the *hidden talents approach*, which proposes that some abilities may be enhanced by adversity, especially skills and abilities that help youth meet real-world challenges in adverse environments [@ellis2017; @ellis2020; @frankenhuis2020; @frankenhuis2013]. These stress-adapted skills are referred to in this literature as *hidden talents*.

The hidden talents approach is rooted in an adaptation-based framework [see @ellis2017; @ellis2020; @frankenhuis2020; @frankenhuis2013 for reviews], which assumes development tailors skills and abilities in relation to the environment. Accordingly, exposure to adversity should shape skills and abilities well-suited for dealing with the challenges and opportunities associated with harsh and unpredictable environments. However, because specific dimensions of adversity (e.g., violence versus resource-scarcity) may pose unique challenges [@ellis2020; @frankenhuis2020; @mclaughlin2019; @sheridan2014], the hidden talents approach proposes that cognitive skills should match specific dimensions of adversity [@frankenhuis2020]. For instance, some evidence suggests that physically abused youth can detect angry facial expressions quickly and accurately [@gibb2009; @pollak2008; @pollak2009], people living in poverty have more accurate social perception [@bjornsdottir2017; @dietze2021; @piff2018], and that anxiously attached people can more accurately detect deception [@eindor2014; @shoda2013].

Although some abilities themselves are more ecologically relevant to specific dimensions of adversity, hidden talents may also emerge under ecologically relevant conditions. For example, previous studies of hidden talents have tested cognition under ecologically relevant *contexts and situations*. In particular, studies have experimentally manipulated the current context to be uncertain and mildly stressful (or not) and measured performance on executive functioning and working memory tasks [@mittal2015; @young2018]. Strikingly, people from unpredictable environments showed more efficient switching on attention shifting tasks and more accurate responses on working memory updating tasks, but this effect only emerged under conditions of uncertainty. There was no performance increase under control conditions. However, some studies have found similar patterns without uncertainty [@fields2021; @nweze2021].

Hidden talents may also emerge when testing *content and stimuli* are more ecologically relevant and concrete. Typical cognitive tests use abstract content that is largely detached from the real world. Such abstract materials may disadvantage youth whose childhood experiences involve less abstract, analytical problem solving [@ellis2020; @frankenhuis2013]. Cultural psychologists and anthropologists have demonstrated this fact across the world. For example, people from non-Western cultures tend to score lower than people in Western cultures on standard versions of cognitive tests, but this performance gap closes or disappears when more ecologically relevant measures are used [@banerjee2017; @grigorenko2004; @rogoff2017; @sternberg2001]. This work challenges the assumption that "baseline" performance is best captured by standard/abstract tests. In fact, they suggest that abstract stimuli, and more broadly standard testing conditions, are specific contexts that may not match one's lived experience [@doebel2020; @zuilkowski2016]. These ideas also raise an important question: do adversity-exposed youth score higher on tests with ecologically-relevant stimuli relative to those with abstract stimuli?

One way to test this question is to replace sterile testing materials with ecologically relevant content from the real world. For example, one study tested memory and reasoning among young adults who varied in their exposures to violence [@frankenhuis2020a]. They created two versions of a memory and reasoning task, one neutral (chronological age) and one more ecologically relevant (social dominance). They found that people with more *current* exposures to violence were equally, or more, accurate when memorizing social-dominance relationships than people with fewer exposures. Notably, such violence exposures were associated with less accurate memory for neutral (age) relationships. This pattern was specific to *current* exposures: people with more *childhood* exposures to violence memorized both content types less accurately. There were no associations between violence exposures and reasoning performance. Overall, this study suggests that ecologically relevant testing materials may in some conditions 'level the playing field' for adversity-exposed youth [@ellis2020; @frankenhuis2020; @rifkin-graboi2021; @vantassel-baska2018].

Although research testing performance in ecologically relevant contexts/situations and using ecologically relevant materials have generated innovative ideas and findings, it remains unclear whether or how these ideas and findings can be practically applied. For example, narrow abilities, such as memorizing social dominance hierarchies or accurately detecting anger in faces, seem to have limited practical value in educational contexts. In addition, although executive functions and working memory are related to many educational outcomes, it is unclear how schools and teachers might leverage skills and abilities that specifically depend on inducing uncertainty. That seems impractical, at best, and unethical, at worst.

Here we address this issue by evaluating the role of ecologically relevant content in attention-shifting and working memory tasks. We define ecologically relevant content as testing material that resembles the local environment and therefore has more meaning and consequence. We used two criteria for stimulus selection. First, we selected stimuli that youth would encounter and interact with on a daily basis. Second, we selected stimuli believed to be more consequential or salient to adversity-exposed youth. For example, money is more consequential and perhaps salient to adversity-exposed youth than other neutral objects (a book, dishes, etc.). We target both within-subject and between-subject patterns of improvement. That is, improvements in performance relative to oneself (within-subjects -- one's performance on abstract compared to ecological task versions), relative to others (between-subjects -- high versus low adversity), or a combination, are theoretically and practically important. In the current study, we expected all youth would score higher on tasks with ecological stimuli. However, we expected this increase to be larger among youth living in adverse conditions compared to youth in supportive conditions.

@@

We then tested performance on two widely used tasks: attention shifting and working memory updating. There were two versions of each task. The first used standard abstract stimuli and the second replaced these stimuli with more ecological stimuli from the real-world (see Figure 1). We then tested the interactive effect of task content and adversity exposure. As such, our primary analyses involved two preregistered confirmatory analyses and one exploratory analysis. Our confirmatory analyses focused on environmental unpredictability and violence exposure. Both dimensions have been associated with improved performance under ecologically relevant conditions (contexts and testing materials) and represent specific environmental challenges. Our exploratory analyses focused on exposure to poverty. Poverty is both relevant to hidden talents (e.g., hidden talents could develop in response to specific challenges of poverty) and a general proxy for adversity rather than a specific exposure.

@@

The current study collected data from `r txt1_sam$raw$n$total` youth (*M~age~* = `r txt1_sam$raw$age$mean`, *SD~age~* = `r txt1_sam$raw$age$sd`) from two sources across Salt Lake City, Utah. The first was a middle school and the second were 5 after-school clubs. The middle school sample (seventh and eighth graders, *N* = `r txt1_sam$raw$n$wjms`; `r txt1_sam$raw$sex$wjms$females` females) formed `r (txt1_sam$raw$n$wjms/txt1_sam$raw$n$total * 100) %>% round(2)`% of the total sample and is socioeconomically and ethnically diverse: `r data1_raw %>% filter(race == 1) %>% nrow()` youth identified as Asian, `r data1_raw %>% filter(race == 2) %>%nrow()` as Native American, `r data1_raw %>% filter(race == 3) %>% nrow()` as African American, `r data1_raw %>% filter(race == 4) %>% nrow()` as White, `r data1_raw %>% filter(race == 5) %>% nrow()` as Hispanic, `r data1_raw %>% filter(race == 6)  %>% nrow()` as Pacific Islander, `r data1_raw %>% filter(race == 7) %>% nrow()` with multiple racial identities, and `r data1_raw %>% filter(is.na(race)) %>% nrow()` were missing. According to school district records, `r (txt1_sam$raw$free_lunch$wjms/txt1_sam$raw$n$wjms * 100) %>% round(2)`% of students were receiving economic assistance based on family income and/or homelessness (i.e., free or reduced-price lunch or fee waivers). The club sample (*N* = `r txt1_sam$raw$n$bgc`; `r txt1_sam$raw$sex$bgc$female` females) contained a wider age range (*M~age~*= `r txt1_sam$raw$age$bgc$mean`, *SD~age~* = `r txt1_sam$raw$age$bgc$sd`) and were also ethnically diverse. `r (txt1_sam$raw$free_lunch$bgc/txt1_sam$raw$n$bgc * 100) %>% round(2)`% were receiving free or reduced-price lunch.

@@

Our measure of environmental unpredictability integrated information from the interview and questionnaire. The interview contained three variables of interest. The first measured the level of exposure to living in a biologically non-intact family from birth to age 13. Interviewers asked participants about their family composition and the adults that lived with them in the home. If a participant's family composition changed, for example, because of a divorce, interviewers recorded the ages when participants experienced the change and for how long it persisted. A family was considered non-intact if the participant did not live with both biological parents during any period from birth to age 13. Based on the earliest age of disruption, we indexed the onset of exposure to a non-intact family by creating a four-level variable from 0 to 3. A 0 indicated the participant lived in an intact family (lived with both biological parents) from birth until age 13 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 0) %>% sum(na.rm=T)`); a 1 indicated that a disruption occurred after the age of 5 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 1) %>% sum(na.rm=T)`); a 2 indicated that a disruption occurred before age 5 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 2) %>% sum(na.rm=T)`); and a 3 indicated the participant was born into a non-intact family (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 3) %>% sum(na.rm=T)`). `r is.na(data2_non_arbitrary$unp_interview_disrupted) %>% sum(na.rm=T) %>% english::as.english() %>% str_to_title()` participants had missing information.

@@

We assessed each participant's poverty exposure using three information sources: interview-assessed parent education and occupational prestige, school or club-provided economic information, and self-reported resource scarcity.

@@

```{r table2, tab.id = "table2", tab.cap.style = "Table Caption", tab.cap = "Median standardized effects, 95% confidence intervals, percent significant, and bootstrapped p-values for the main effect of adversity and adversity by task version interactions."}
table.02.0 %>% 
  ungroup() %>% 
  mutate(iv = factor(iv, c("Unpredictability","Violence","Poverty"))) %>%
  arrange(dv,iv) %>% 
  select(-dv) %>% 
  add_column(.after = 5,"empty1" = "") %>% 
  add_row(.before = 1, "iv" = "Attention Shifting") %>% 
  add_row(.after = 4, "iv" = "Working Memory Updating") %>% 
  flextable() %>% 
  merge_h_range(i = 1, j1 = 1, j2 = 10) %>% 
  merge_h_range(i = 5, j1 = 1, j2 = 10) %>% 
  bold(i = c(1,5), j = 1) %>% 
  padding(i = c(2:4, 6:8), j = 1, padding.left = 10) %>% 
  width("empty1", width = .2) %>% 
  set_header_labels(iv = "", empty1 = "", ci_main = "95% CI", ci_int = "95% CI") %>% 
  add_header_row(
    values = c(" ", "Main Effect", " ", "Interaction"),
    colwidths = c(1, 4, 1, 4)
  ) %>% 
  compose(i = 2, j = c(2,7), as_paragraph("\U1D6FD"), part = "header") %>% 
  compose(i = 2, j = c(4,9), as_paragraph(as_i("p "), "(%)"), part = "header") %>% 
  compose(i = 2, j = c(5,10), as_paragraph(as_i("p")), part = "header") %>% 
  align(i = 1:2, align = "center", part = "header") %>% 
  border_remove() %>% 
  border(i = 1, border.top = fp_border_default(), part = "header") %>% 
  border(i = 1, j = c(2:5,7:10), border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 2, border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 8, border.bottom = fp_border_default(), part = "body") %>% 
  bold(i = 1:2, part = "header") %>% 
  set_table_properties(width = 1, layout = "autofit") %>% 
  add_footer_row(
    values = " ",
    colwidths = 10
  ) %>% 
    add_footer_row(
    values = " ",
    colwidths = 10
  ) %>% 
  compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "The p (%) column reflects the number of analyses that produced p-values < .05 for a given multiverse. The total number of analyses for each measure was 64. Overall p-values were computed using a bootstrapped resampling method and reflect the probability of obtaining an effect size as extreme or more extreme given the median effect is 0."), 
    part = "footer"
  )
```

@@

```{r table3, tab.id = "table3", tab.cap.style = "Table Caption", tab.cap = "Median simple effects for task version and adversity."}
table.03.0 %>% 
  ungroup() %>%
  mutate(iv = factor(iv, c("Unpredictability","Violence","Poverty"))) %>%
  arrange(dv,iv) %>% 
  select(-dv) %>% 
  add_column(.after = 5,"empty1" = "") %>% 
  add_row(.before = 1, "iv" = "Attention Shifting") %>% 
  add_row(.after = 4, "iv" = "Working Memory Updating") %>% 
  flextable() %>% 
  merge_h_range(i = 1, j1 = 1, j2 = 10) %>% 
  merge_h_range(i = 5, j1 = 1, j2 = 10) %>% 
  bold(i = c(1,5), j = 1) %>% 
  padding(i = c(2:4, 6:8), j = 1, padding.left = 10) %>% 
  width("empty1", width = .2) %>% 
  set_header_labels(iv = "", empty1 = "") %>% 
  add_header_row(
    values = c(" ", "Task Version", " ", "Adversity Exposure"),
    colwidths = c(1, 4, 1, 4)
  ) %>% 
  compose(i = 2, j = c(2), as_paragraph(as_i("b "), "(abstract)"), part = "header") %>% 
  compose(i = 2, j = c(4), as_paragraph(as_i("b "), "(ecological)"), part = "header") %>% 
  compose(i = 2, j = c(7), as_paragraph(as_i("b "), "(low)"), part = "header") %>% 
  compose(i = 2, j = c(9), as_paragraph(as_i("b "), "(high)"), part = "header") %>% 
  compose(i = 2, j = c(3,8), as_paragraph(as_i("p "), "(%)"), part = "header") %>% 
  compose(i = 2, j = c(5,10), as_paragraph(as_i("p "), "(%)"), part = "header") %>% 
  align(i = 1:2, align = "center", part = "header") %>% 
  border_remove() %>% 
  border(i = 1, border.top = fp_border_default(), part = "header") %>% 
  border(i = 1, j = c(2:5,7:10), border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 2, border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 8, border.bottom = fp_border_default(), part = "body") %>% 
  bold(i = 1:2, part = "header") %>% 
  fontsize(size = 11, part = "body") %>%
  set_table_properties(width = 1, layout = "autofit") %>% 
  add_footer_row(
    values = " ",
    colwidths = 10
  ) %>% 
    add_footer_row(
    values = " ",
    colwidths = 10
  ) %>% 
  compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "Simple effects are unstandardized. Task Version reflects the median simple effect of adversity when the task version was abstract and ecological. Adversity reflects the median simple effect of task version when adversity is low (-1 SD) and high (+1 SD). The p (%) columns reflect the number of analyses that produced simple effect p-values < .05."), 
    part = "footer"
  )
```

@@

For this expanded multiverse approach, we retained all original data processing decisions (i.e., original 64 combinations of decisions) and added adversity component as an additional decision. For unpredictability, we analyzed each unpredictability component and the composite measure (composite analyses are redundant with primary analyses but were retained for comparison purposes), which resulted in 384 analyses (64 original × three versions of unpredictability × two performance measures: attention shifting and working memory). We applied the same approach to violence exposure (three violence variables, 384 analyses) and poverty exposure (four poverty variables, 512 analyses). The full results of these analyses are reported in the supplement (see SI Table 2).

@@

Leveraging multiverse analysis [@simonsohn2020; @steegen2016], we tested interactions between dimensions of adversity and task version on attention shifting and working memory updating performance. We did not find interaction effects for attention shifting. However, youth exposed to unpredictability and violence showed lower switch costs on average, suggesting slightly improved attention shifting performance irrespective of task content. These findings are consistent with a recent study showing that higher caregiving instability was associated with improved attention shifting [@fields2021]. However, our findings also partially diverge from prior work [@mittal2015; @nweze2021]. For example, whereas we found unpredictability was associated with improved attention shifting on average, Mittal and colleagues [-@mittal2015] found no difference in performance under neutral conditions. And, although they found improved attention shifting for those exposed to high unpredictability under uncertainty, it remains unclear how this effect relates to those obtained in the current study. In addition, Nweze and colleagues [-@nweze2021] found no association between adversity and attention shifting, but they tested institutionalized children rather than measuring adversity dimensions. In general, we should interpret the results cautiously. Although we followed field standards for calculating switch costs by computing difference scores, such scores ignore participant-level reaction times. Mixed-modeling approaches can account for both overall reaction times and differences across switch and repeat trials, but require more statistical power to test three-way interactions. Future research should consider modeling approaches that capture participant-level reaction times and differences across conditions.

@@

Overall, inconsistencies across prior work and the current attention-shifting and working memory updating findings could be attributed to differences in sampling (adults vs. youth), the sampled range of adversity exposure (restricted vs. broad), or methods/design (within- vs. between-subjects design; lab vs. community/school settings). Future research should both address these inconsistencies and conduct more direct replications of individual studies to converge on reliable patterns.
